% LaTeX document with cross-references requiring 2 passes
% Tests multi-pass compilation for reference resolution
\documentclass{article}

\title{Cross-Reference Test Document}
\author{Test Suite}
\date{\today}

\begin{document}

\maketitle

\section{Introduction}
\label{sec:intro}

This document tests cross-reference resolution which requires multiple
compilation passes. On the first pass, references appear as ``??'', and
only on the second pass do they resolve to actual numbers.

See Section~\ref{sec:methodology} for details on the methodology.

\section{Methodology}
\label{sec:methodology}

As mentioned in Section~\ref{sec:intro}, this document requires two passes.
The methodology is described in detail below.

Consider Equation~\ref{eq:test} which demonstrates a simple formula.

\begin{equation}
  \label{eq:test}
  a^2 + b^2 = c^2
\end{equation}

\subsection{Test Cases}
\label{subsec:cases}

The test cases in Subsection~\ref{subsec:cases} verify that all
references resolve correctly on the second pass.

\section{Results}
\label{sec:results}

Results are summarized in Section~\ref{sec:results} (self-reference test).
For background, see Section~\ref{sec:intro}.

\subsection{Forward References}

This subsection tests forward references to Section~\ref{sec:conclusion}
which appears later in the document.

\section{Conclusion}
\label{sec:conclusion}

This document validates multi-pass compilation. All references should
resolve correctly, including:

\begin{itemize}
  \item Section~\ref{sec:intro}
  \item Section~\ref{sec:methodology}
  \item Subsection~\ref{subsec:cases}
  \item Section~\ref{sec:results}
  \item Equation~\ref{eq:test}
\end{itemize}

\end{document}
